---
title: "lab6"
author: "Joanna Tang"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)


penguins <- read_csv("penguins.csv")
```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

Since we're looking at 1 factor (species), but have >2 samples/levels (species), we know we'll be doing ANOVA.  First, we have to fulfill ANOVA assumptions.

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex)) # Create table with penguins data, counting how many of each species and sex

counts

penguins2 <- penguins %>%
  mutate(common_name = case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
    species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
    species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
  )) %>% # Shortening species name
  select(-species) %>% # Remove species column
  filter(sex == "FEMALE") # Retain only females

penguins2
```

- Exploratory graphs
- Summary statistics
- Levene's test for equal variances (analagous to 2-sample f-test)
- one-way ANOVA
- post-hoc testing

```{r}

penguins_hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) + # Color bars differently for each common_name
  facet_wrap(~common_name) # Separates each species into its own graph

penguins_hists


penguins_qqplots <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~common_name) # Making side-by-side qqplots for each species

penguins_qqplots


# Levene's test for equal variances (>2 groups...if you have 2 groups, use the F-test for equal variances)

# We'll use the function leveneTest from the "car" package

# H0: Variances equal
# HA: Variances unequal

penguins_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)

penguins_levene
# p = .058, we retain the null hypothesis of equal variances


penguins_variances <- penguins2 %>%
  group_by(common_name) %>% 
                              summarize(
                                variance = var(flipper_length)
                              )

penguins_variances
# largest variances is less than 4x smallest variance --> satisfies ANOVA assumptions (even if I got a significant result from Levene's Test)

```

#3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins?

```{r}
# H0: Mean flipper lengths across all species are equal
# HA: AT LEAST 2 means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)
# p<.01 --> If these samples are from populations that actually have equal means, then there is a very unlikely chance that we would have taken samples with means at least this different by random chance.
# I conclude that at least 2 samples were taken from populations with different means.  There is a significant difference in mean flipper length between penguin species.
# Which ones are different?  Are all 3 different from each other?

#Post-hoc testing using Tukey's HSD
penguin_post_hoc <- TukeyHSD(penguin_aov)
penguin_post_hoc

```
Mean flipper lenghts for Adelie, Chinstrap, and Gentoo penguins are all significantly different from each other (F(df) = , p<.001, $\alpha$ = .05, with post-hoc Tukey's HSD, $\alpha$ = .05)
