---
title: "lab6"
author: "Joanna Tang"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)


penguins <- read_csv("penguins.csv")
```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

Since we're looking at 1 factor (flipper length), but have >2 samples/levels (species), we know we'll be doing ANOVA.  First, we have to fulfill ANOVA assumptions.

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex)) # Create table with penguins data, counting how many of each species and sex

counts

penguins2 <- penguins %>%
  mutate(common_name = case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
    species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
    species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
  )) %>% # Shortening species name
  select(-species) %>% # Remove species column
  filter(sex == "FEMALE") # Retain only females

penguins2
```

- Exploratory graphs
- Summary statistics
- Levene's test for equal variances (analagous to 2-sample f-test)
- one-way ANOVA
- post-hoc testing